---
title: "Prediction with cluster representatives"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction with cluster representatives}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The getting-started guide (the package [home page](https://gregfaletto.github.io/cssr-project/)) shows how to *select* features and clusters with cluster stability selection. This vignette shows how to use a fitted model to **predict** the response for new observations.

Cluster stability selection predicts by forming **cluster representatives** — weighted averages of the members of each selected cluster — and fitting ordinary least squares on those representatives. Because each cluster member is a noisy proxy for the same underlying signal, averaging the members estimates that signal better than any single member would, which tends to improve prediction. (See [Faletto and Bien, 2022](https://arxiv.org/abs/2201.00494).)

```{r setup}
library(cssr)
set.seed(983219)

# Data with a cluster of 10 correlated proxies (features 1-10), split into a
# training set and a test set.
data <- genClusteredData(n = 120, p = 40, cluster_size = 10,
                         k_unclustered = 10, snr = 3)

X_train <- data$X[1:80, ]
y_train <- data$y[1:80]
X_test  <- data$X[81:120, ]
y_test  <- data$y[81:120]

clusters <- list("Z_cluster" = 1:10)
```

## The quick way: `cssPredict()`

`cssPredict()` runs the whole pipeline — selection, cluster-representative formation, the downstream regression, and prediction — in a single call. It is the easiest way to get predictions.

```{r}
preds <- cssPredict(X_train, y_train, X_test, clusters)
head(preds)

# Test-set root mean squared error
sqrt(mean((preds - y_test)^2))
```

## More control: `css()` + `getCssPreds()`

As in the getting-started guide, `css()` does the expensive stability-selection step once; you can then reuse its output. To predict, `getCssPreds()` forms the cluster representatives and fits the downstream ordinary-least-squares model.

That regression needs its own training data, kept separate from the data used for selection. The simplest way is to reserve some rows with `css()`'s `train_inds` argument — those rows are held out of the selection step and used only to fit the downstream model.

```{r}
lambda <- getLassoLambda(X_train, y_train)

# Reserve rows 1-40 to fit the downstream model; the rest drive selection.
css_output <- css(X_train, y_train, lambda, clusters = clusters,
                  train_inds = 1:40)

preds_full <- getCssPreds(css_output, testX = X_test,
                          weighting = "weighted_avg", cutoff = 0.3)
head(preds_full)
```

Alternatively, if you ran `css()` without `train_inds`, you can supply the downstream training data directly through `trainX`/`trainY`. As with `train_inds`, this should ideally be data not used in the selection step (here we reuse `X_train` just to keep the example short):

```{r}
css_no_train <- css(X_train, y_train, lambda, clusters = clusters)

preds_ext <- getCssPreds(css_no_train, testX = X_test, trainX = X_train,
                         trainY = y_train, weighting = "weighted_avg",
                         cutoff = 0.3)
head(preds_ext)
```

## Weighting schemes

The `weighting` argument controls how cluster representatives are formed:

* `"simple_avg"` — a plain (equally weighted) average of the cluster members. Optimal when the proxies are equally noisy.
* `"weighted_avg"` — a weighted average with weights inferred from the selection proportions, so noisier proxies get less weight. Best when the members' noise levels vary.
* `"sparse"` — use a single member (the most-selected one), in the spirit of the protolasso. Simplest, but discards the other members.

```{r}
for (w in c("simple_avg", "weighted_avg", "sparse")) {
  p <- getCssPreds(css_output, testX = X_test, weighting = w, cutoff = 0.3)
  cat(sprintf("%-12s test RMSE: %.3f\n", w, sqrt(mean((p - y_test)^2))))
}
```

## Just the design matrix: `getCssDesign()`

If you would rather build your own downstream model (anything other than OLS), `getCssDesign()` returns the matrix of cluster representatives for new data, which you can pass to any modeling function.

```{r}
X_test_reps <- getCssDesign(css_output, newX = X_test,
                            weighting = "weighted_avg", cutoff = 0.3)

# One column per selected cluster
dim(X_test_reps)
head(X_test_reps)
```

## See also

* The [getting-started guide](https://gregfaletto.github.io/cssr-project/) for selecting features and clusters.
* `vignette("advanced-usage", "cssr")` for competitor methods, the data generators, and other helpers.
