---
title: "Advanced usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette covers additional features of the package: the competitor methods compared against in the paper, helpers for choosing a model size, and the data generators used in the simulation studies. For the core workflow, see the [getting-started guide](https://gregfaletto.github.io/cssr-project/) (which also covers custom base selectors via `fitfun` and choosing `lambda` with `getLassoLambda()`) and `vignette("prediction", "cssr")`.

```{r setup}
library(cssr)
set.seed(983219)

data <- genClusteredData(n = 80, p = 40, cluster_size = 10,
                         k_unclustered = 10, snr = 3)
X <- data$X
y <- data$y
clusters <- list("Z_cluster" = 1:10)
```

## Competitor methods

The package ships the two competitor methods that [Faletto and Bien (2022)](https://arxiv.org/abs/2201.00494) compare against. Both handle clusters of correlated features, but differently:

* `protolasso()` ([Reid and Tibshirani, 2016](https://doi.org/10.1093/biostatistics/kxv049)) picks a single *prototype* from each cluster and then runs the lasso.
* `clusterRepLasso()` ([Bühlmann et al., 2013](https://doi.org/10.1016/j.jspi.2013.05.019)) replaces each cluster with its *average* (a cluster representative) and then runs the lasso.

```{r}
proto <- protolasso(X, y, clusters)
str(proto, max.level = 1)

crl <- clusterRepLasso(X, y, clusters)
str(crl, max.level = 1)
```

Both return selected sets along a lasso path. In the paper's simulations and data application, cluster stability selection is generally more stable and predictive than either; see `vignette("prediction", "cssr")` and the paper for the comparison.

## Choosing a model size: `getModelSize()`

`getModelSize()` suggests a data-driven model size by fitting the lasso on the full data set. It is a convenient default for the `max_num_clusts` argument of `cssSelect()` / `getCssSelections()`.

```{r}
getModelSize(X, y, clusters)
```

## Data generators for simulations

`genClusteredData()` (used above and in the getting-started guide) is the basic generator. Two richer variants reproduce the settings in the paper's simulation studies, where cluster members are proxies of *differing* quality:

* `genClusteredDataWeighted()` — cluster members have different (fixed) noise levels; `n_strong_cluster_vars` of them are low-noise "strong" proxies.
* `genClusteredDataWeightedRandom()` — like the above, but the members' noise levels are drawn at random.

```{r}
set.seed(1)
dw <- genClusteredDataWeighted(n = 80, p = 40, k_unclustered = 10,
                               cluster_size = 10, n_strong_cluster_vars = 3,
                               snr = 3)
str(dw, max.level = 1)

dwr <- genClusteredDataWeightedRandom(n = 80, p = 40, k_unclustered = 10,
                                      cluster_size = 10, snr = 3)
str(dwr, max.level = 1)
```

The errors-in-variables model behind these generators treats each observed proxy as a latent signal plus noise. `getNoiseVar()` returns the noise variance that produces a target correlation between a proxy and the latent signal:

```{r}
# Noise variance giving a proxy-to-latent correlation of 0.9
getNoiseVar(0.9)
```

## See also

* The [getting-started guide](https://gregfaletto.github.io/cssr-project/): selecting features, custom `fitfun`, `getLassoLambda()`.
* `vignette("prediction", "cssr")`: prediction with cluster representatives.
* The [annotated source bookdown](https://gregfaletto.github.io/cssr-project/create/) for the full implementation.
